The imprint of global magma oceans on exoplanet demographics

### 2023-02-10 src/tex/ms.tex
increase sample size of optimistic case

\subsection{Testability of the runaway greenhouse hypothesis}
\begin{figure}[ht!]
%    \script{figurescript.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/optimistic_R-S.pdf}
        \caption{
        Detection of the runaway greenhouse threshold in the optimistic case.
        From simulated radius and instellation measurements of a large ($N = \Nplanets$) survey, we compute the moving average (blue confidence intervals) and fit the runaway greenhouse hypothesis to it (Eqn.~\ref{eq:rgh_hypo}, random draws from the posterior in red).
            The pattern is detected with high significance.
        }
        \label{fig:optimistic_R-S}
    \end{centering}
\end{figure}
Figure~\ref{fig:optimistic_R-S} shows a prototypical detection of the runaway greenhouse threshold.
For the sake of clarity, we chose an optimistic scenario where the sample is large ($N = \Nplanets$) and the measurement uncertainties are small ($\sigma_{R} = \sigmaR, \sigma_{S} = \sigmaS$).
We assumed that the fraction of those planets irradiated stronger than $S_\mathrm{thresh}$ that have runaway greenhouse climates is \frgh, and we chose a water mass fraction of \wrr\ for each planet.
In this case, the pattern was detected with high significance ($\Delta \ln \mathcal{Z} \approx 100$).

\begin{figure}[ht!]
%    \script{figurescript.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/corner.pdf}
        \caption{
            Retrieved posterior distribution of key parameters in the optimistic scenario. The density maps in each panel show relationships between and marginalized distributions of the model parameters as they could be retrieved with a high-precision transit survey and a sample of \Nplanets\ planets. True values of the parameters for the injected effect are shown in orange. The threshold instellation can be well constrained; the predominant water fraction and the fraction of planets with runaway greenhouse climates are degenerate.
        }
        \label{fig:cornerplot}
    \end{centering}
\end{figure}
With such a strong signal, we can attempt an inference of the parameters defining the injected effect.
Figure~\ref{fig:cornerplot} shows the posterior distributions of $S_\mathrm{thresh}$, $x_{H_2O}$, and $f_\mathrm{rgh}$ as determined by the nested sampler.
The threshold instellation can be accurately constrained, $S_\mathrm{thresh} = \var{retrSthresh}$.
Both a higher water mass fraction and a higher fraction of planets in a runaway greenhouse state lead to larger average radii, thus these parameters are strongly correlated and cannot be well constrained without independent measurements.

\begin{note}
    Given an optimistic survey design, what influences the probability of correctly rejecting a false null hypothesis?
    \begin{figure}[ht!]
%    \script{figurescript.py}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/optimistic_statpwr_H2O-f.pdf}
            \caption{
                Statistical power of the runaway greenhouse hypothesis test as a function of model parameters.
                For a sample size $N = \Nplanets$, the color code shows the fraction of simulations resulting in a sound detection ($\Delta \ln \mathcal{Z} > 3$) for different combinations of water mass fraction and fraction of greenhouse climate planets.
                Higher values in either parameter result in a more reliable detection.
            }
            \label{fig:statpwr_H2O-f}
        \end{centering}
    \end{figure}
    Figure~\ref{fig:statpwr_H2O-f} explores the statistical power achieved in the above scenario for different combinations of the poorly constrained parameters $x_{H_2O}$ and $f_\mathrm{rgh}$.
    The higher the water content and thus the radius change, and the higher the fraction of planets in runaway greenhouse climates, the greater the prospect of a statistically sound detection.
     $f_\mathrm{rgh}$ dominates this trend: It enters linearly into the average planet radius, whereas the contribution of $x_{H_2O}$ - as predicted by the geophysical models - is sublinear with a power-law exponent of $\sim 0.3$.
    As long as $f_\mathrm{rgh}$ is larger than $\sim 0.1$, a sample size of \Nplanets\ is sufficient for a $\SI{50}{\percent}$ detection rate even for water ratios as low as $x_{H_2O} = 0.001$.
\end{note}

>>>>

\subsection{Testability of the runaway greenhouse hypothesis}
\begin{figure}[ht!]
%    \script{figurescript.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/optimistic_R-S.pdf}
        \caption{
        Detection of the runaway greenhouse threshold in the optimistic case.
        From simulated radius and instellation measurements of a large ($N = \var{N_optimistic}$) survey, we compute the moving average (blue confidence intervals) and fit the runaway greenhouse hypothesis to it (Eqn.~\ref{eq:rgh_hypo}, random draws from the posterior in red).
            The pattern is detected with high significance.
        }
        \label{fig:optimistic_R-S}
    \end{centering}
\end{figure}
Figure~\ref{fig:optimistic_R-S} shows a prototypical detection of the runaway greenhouse threshold.
For the sake of clarity, we chose an optimistic scenario where the sample is large ($N = \var{N_optimistic}$) and the measurement uncertainties are small ($\sigma_{R} = \sigmaR, \sigma_{S} = \sigmaS$).
We assumed that the fraction of those planets irradiated stronger than $S_\mathrm{thresh}$ that have runaway greenhouse climates is \frgh, and we chose a water mass fraction of \wrr\ for each planet.
In this case, the pattern was detected with high significance ($\Delta \ln \mathcal{Z} \approx 100$).

\begin{figure}[ht!]
%    \script{figurescript.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/corner.pdf}
        \caption{
            Retrieved posterior distribution of key parameters in the optimistic scenario. The density maps in each panel show relationships between and marginalized distributions of the model parameters as they could be retrieved with a high-precision transit survey and a sample of \var{N_optimistic} planets. True values of the parameters for the injected effect are shown in orange. The threshold instellation can be well constrained; the predominant water fraction and the fraction of planets with runaway greenhouse climates are degenerate.
        }
        \label{fig:cornerplot}
    \end{centering}
\end{figure}
With such a strong signal, we can attempt an inference of the parameters defining the injected effect.
Figure~\ref{fig:cornerplot} shows the posterior distributions of $S_\mathrm{thresh}$, $x_{H_2O}$, and $f_\mathrm{rgh}$ as determined by the nested sampler.
The threshold instellation can be accurately constrained with a median $\hat{S}_\mathrm{thresh} = \var{retrSthresh}$.
Both a higher water mass fraction and a higher fraction of planets in a runaway greenhouse state lead to larger average radii, thus these parameters are strongly correlated and cannot be well constrained without independent measurements.

\begin{note}
    Given an optimistic survey design, what influences the probability of correctly rejecting a false null hypothesis?
    \begin{figure}[ht!]
%    \script{figurescript.py}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/optimistic_statpwr_H2O-f.pdf}
            \caption{
                Statistical power of the runaway greenhouse hypothesis test as a function of model parameters.
                For a sample size $N = \var{N_optimistic}$, the color code shows the fraction of simulations resulting in a sound detection ($\Delta \ln \mathcal{Z} > 3$) for different combinations of water mass fraction and fraction of greenhouse climate planets.
                Higher values in either parameter result in a more reliable detection.
            }
            \label{fig:statpwr_H2O-f}
        \end{centering}
    \end{figure}
    Figure~\ref{fig:statpwr_H2O-f} explores the statistical power achieved in the above scenario for different combinations of the poorly constrained parameters $x_{H_2O}$ and $f_\mathrm{rgh}$.
    The higher the water content and thus the radius change, and the higher the fraction of planets in runaway greenhouse climates, the greater the prospect of a statistically sound detection.
     $f_\mathrm{rgh}$ dominates this trend: It enters linearly into the average planet radius, whereas the contribution of $x_{H_2O}$ - as predicted by the geophysical models - is sublinear with a power-law exponent of $\sim 0.3$.
    As long as $f_\mathrm{rgh}$ is larger than $\sim 0.1$, a sample size of \var{N_optimistic} is sufficient for a $\SI{50}{\percent}$ detection rate even for water ratios as low as $x_{H_2O} = 0.001$.
\end{note}
